---
title: "MohiniQ1"
author: "Mohini Agarwal"
date: "8/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
library(mosaic)
library(tidyverse)
green = read.csv('../data/greenbuildings.csv')
attach(green)
cols = c("renovated", "class_a", "class_b", "LEED", "Energystar", "green_rating", "net", "amenities")
green[, cols] = data.frame(apply(green[, cols], 2, as.factor))
green$cluster = as.factor(green$cluster)
```

## Median rent of green buildings is higher and overall rent is less spread than other buildings

```{r}
ggplot(data = green) + 
  geom_boxplot(mapping = aes(x = green_rating, y = Rent, fill = green_rating))
```

## Increase in rent with size is more or less the same for both

```{r}
ggplot(data = green) +
  geom_point(mapping = aes(x = size, y = Rent, col = green_rating)) +
  facet_wrap(~ green_rating, nrow = 1)
```

## Confirming the median rent values

```{r}
green %>%
  group_by(green_rating) %>%
  summarize(medianRent = median(Rent, na.rm = TRUE))
```

## Looking at spread of employment growth for both types of buildings

```{r}
ggplot(data = green) + 
  geom_boxplot(mapping = aes(x = green_rating, y = empl_gr, fill = green_rating))
```

## Increase in rent with leasing rate is same for both types of buildings

```{r}
ggplot(data = green) +
  geom_point(mapping = aes(x = leasing_rate, y = Rent, col = green_rating)) +
  facet_wrap(~ green_rating, nrow = 1)
```

## What are the spread of # of stories for both kinds of buildings?

```{r}
ggplot(data = green) + 
  geom_histogram(mapping = aes(x = stories, fill = green_rating), bins = 15) +
  facet_wrap(~green_rating, nrow = 1)
```

## What are the spread of # of stories for both kinds of buildings?

```{r}
ggplot(data = green) + 
  geom_boxplot(mapping = aes(x = green_rating, y = stories, fill = green_rating))
```

