
```{r, warning=FALSE, message = FALSE}
theme_set(theme_classic())
g1 <- ggplot(ABIA, aes(x=(UniqueCarrier), ArrDelay))
g1 + geom_boxplot(fill="plum", outlier.shape = NA) + 
  scale_y_continuous(limits = quantile(ABIA$ArrDelay, c(0.1, 0.9))) +
  labs(title="Box plot", 
       subtitle="Arr Delay grouped by Carrier",
       caption="Source: ABIA",
       x="Carrier",
       y="Arrival Delay")
```
Similar to their departure status, most of the carriers's flights arrive on average on or before the scheduled arrival time.

```{r}
ABIA_summ_arr = ABIA %>%
  group_by(UniqueCarrier)  %>%  # group the data points by model name
  summarize(Arr.mean = mean(ArrDelay))  # calculate a mean for each model
```
```{r}
# reorder the x labels
ggplot(ABIA_summ_arr, aes(x=reorder(UniqueCarrier, Arr.mean), y=Arr.mean)) + 
  geom_bar(stat='identity') + 
  coord_flip()
```
```{r}
# first add a new column using the mutate function
ABIA_summ_arr = mutate(ABIA_summ_arr, Arr.z = (Arr.mean - mean(Arr.mean))/sd(Arr.mean))

# now plot the newly defined variable
ggplot(ABIA_summ_arr, aes(x=reorder(UniqueCarrier, Arr.z), y=Arr.z)) + 
  geom_bar(stat='identity') + 
  labs(title="Arr Delay by Carrier", 
       caption="Source: ABIA",
       y="Arr Delay (z-score)",
       x = "Carrier") + 
  coord_flip()
```
This shows that Carrier OH has the highest average arrival delay of any carrier and US has the lowest average arrival delay of any carrier. 

```{r}
#
ABIA$Cancelled = factor(ABIA$Cancelled)
canc_flights = ABIA %>%
  group_by(Cancelled, CancellationCode) %>%
  summarize(CancelledCount = count(FlightNum))

# 
ggplot(data = canc_flights) +
  geom_bar(mapping = aes(x=Cancelled, y = CancelledCount), stat = 'identity', fill = 'plum')+
  facet_wrap(~CancellationCode)
```

This shows that hardly any flights get cancelled that depart from Austin. Out of the ones that get cancelled, most of them are due to weather related reasons.


```{r}
#
ABIA$Diverted = factor(ABIA$Diverted)
div_flights = ABIA %>%
  group_by(Diverted) %>%
  summarize(DivertedCount = count(FlightNum))

# 
ggplot(data = div_flights) +
  geom_bar(mapping = aes(x=Diverted, y = DivertedCount), stat = 'identity', fill = 'plum')
```

This shows that hardly any flights get diverted that depart from Austin.

