---
title: "Visual story telling part 2: flights at ABIA"
output: html_document
---

```{r, warning=FALSE, message = FALSE}
library("tidyverse")
ABIA <- read.csv("C:/Users/rawin/Documents/GitHub/STA380/data/ABIA.csv")
View(ABIA)
ABIA[is.na(ABIA)] = 0
```

```{r}
ABIA$date <- as.Date(with(ABIA, paste(Year, Month, DayofMonth,sep="-")), "%Y-%m-%d")
```

```{r, warning=FALSE, message = FALSE}
theme_set(theme_classic())
g <- ggplot(ABIA, aes(x=(UniqueCarrier), DepDelay))
g + geom_boxplot(fill="plum", outlier.shape = NA) + 
  scale_y_continuous(limits = quantile(ABIA$DepDelay, c(0.1, 0.9))) +
    labs(title="Box plot", 
         subtitle="Dep Delay grouped by Carrier",
         caption="Source: ABIA",
         x="Carrier",
         y="Departure Delay")
```
Most of the carriers's flights depart on average on or before the scheduled departure time. 
```{r, warning=FALSE, message = FALSE}
theme_set(theme_classic())
g <- ggplot(ABIA, aes(UniqueCarrier, DepDelay))
g + geom_violin(fill="plum", outlier.shape = NA) + 
  scale_y_continuous(limits = quantile(ABIA$DepDelay, c(0.1, 0.9))) +
    labs(title="Violin plot", 
         subtitle="Dep Delay grouped by Carrier",
         caption="Source: ABIA",
         x="Carrier",
         y="Departure Delay")
```
The violin plot shows that some carriers are stingy when it comes to when the flight departs. Carriers OH and YV like their flights to depart right on time instead of early.
```{r}
ABIA_summ = ABIA %>%
  group_by(UniqueCarrier)  %>%  # group the data points by model name
  summarize(Del.mean = mean(DepDelay))  # calculate a mean for each model
```
```{r}
# reorder the x labels
ggplot(ABIA_summ, aes(x=reorder(UniqueCarrier, Del.mean), y=Del.mean)) + 
  geom_bar(stat='identity') + 
  coord_flip()
```
```{r}
# first add a new column using the mutate function
ABIA_summ = mutate(ABIA_summ, Del.z = (Del.mean - mean(Del.mean))/sd(Del.mean))

# now plot the newly defined variable
ggplot(ABIA_summ, aes(x=reorder(UniqueCarrier, Del.z), y=Del.z)) + 
  geom_bar(stat='identity') + 
  labs(title="Dep Delay by Carrier", 
       caption="Source: ABIA",
       y="Dep Delay (z-score)",
       x = "Carrier") + 
  coord_flip()
```
This shows that Carrier EV has the highest average departure delay of any carrier and US has the lowest average departure delay of any carrier. 

```{r}
ABIA$delay = names(ABIA[c('CarrierDelay','WeatherDelay','NASDelay','SecurityDelay','LateAircraftDelay')])[max.col(ABIA[c('CarrierDelay','WeatherDelay','NASDelay','SecurityDelay','LateAircraftDelay')])]
ABIA$delay = factor(ABIA$delay)
```

```{r}
ABIA$Month = factor(ABIA$Month)
ABIA$DayOfWeek = factor(ABIA$DayOfWeek)
```

```{r}
ggplot(ABIA,aes(factor(Month))) +
  geom_bar(aes(fill=factor(delay)), position = 'dodge') +
  labs(title="Flight Delays in Austin", 
       y="Number of Delayed Flights",
       x = "Month") + scale_fill_brewer(palette="Dark2")
  
```
There seems to be less overall flight delays during September, October, and November. Most flight delays are caused by late aircraft.

```{r}
ggplot(ABIA,aes(factor(DayOfWeek))) +
  geom_bar(aes(fill=factor(delay)), position = 'dodge') +
  labs(title="Flight Delays in Austin", 
       y="Number of Delayed Flights",
       x = "Day of the Week (1: Monday)") + scale_fill_brewer(palette="Dark2")
```
There seems to be significantly less aircraft delays on Saturdays. Once again, most aircraft delays are due to late aircraft.








