---
title: "Market Segmentation"
output: html_document
---
```{r}
library(ggplot2)
library(LICORS)  # for kmeans++
library(foreach)
library(mosaic)
library(dplyr)
library(plotly)
library(factoextra)
```

```{r}
market = read.csv("C:/Users/rawin/OneDrive/UT AUSTIN MSBA/Predictive Modeling/Part 2 Exercises/Data/social_marketing.csv",row.names = 1, header = TRUE)

```

Pre-process data to by grouping similar variables to reduce number of variables:
```{r}
X = market %>% select(-c(chatter,uncategorized,spam,adult))

#X$media = X$current_events + X$politics + X$news + X$tv_film + X$sports_fandom
#X = X %>% select(-c(current_events,politics,news,tv_film,sports_fandom))

#X$fitness = X$sports_playing + X$personal_fitness
#X = X %>% select(-c(sports_playing,personal_fitness))

#X$education = X$college_uni + X$school
#X = X %>% select(-c(college_uni,school))

#X$artsAndcrafts = X$art + X$crafts
#X = X %>% select(-c(art,crafts))

#X$Business = X$business + X$small_business
#X = X %>% select(-c(business,small_business))

```

Choosing K using Gap Statistic: did not converge
Choosing K using the Elbow Method:
```{r}
X_scaled = scale(X, center=TRUE, scale=TRUE)
k_grid = seq(2,20,by = 1)
SSE_grid = foreach(k = k_grid, .combine = 'c') %do% { 
  cluster_k = kmeans(X_scaled, k, nstart = 50)
  cluster_k$tot.withinss
}
plot(k_grid, SSE_grid)
```

No defiitive elbow - will make an educated guess on k

Clustering using kmeans++ initialization:
```{r}
X_scaled = scale(X, center=TRUE, scale=TRUE)

# Using kmeans++ initialization
#Using k=6
clust_k6 = kmeanspp(X_scaled, k=6, nstart=25)

#Using k=10
clust_k10 = kmeanspp(X_scaled, k=10, nstart=25)

##within cluster sum of squares
sprintf("Within cluster average distance for k=6 is %f", sum(clust_k6$withinss))
sprintf("Within cluster average distance for k=10 is %f", sum(clust_k10$withinss))

##Between cluster distances?
sprintf("Between cluster average distance for k=6 is %f", clust_k6$betweenss)
sprintf("Between cluster average distance for k=10 is %f", clust_k10$betweenss)

```
k=10 has smaller average distance within clusters and larger average distance between clusters. Therefore k=10 is the better k value.

```{r}
for (i in 1:10){
  len = length(which(data.frame(X_scaled, clust_k10$cluster) == i))
  cat("Cluster", i, "has" , len, "members\n")
}
```

```{r}
## get cluster means 
cluster_means = aggregate(X_scaled,by=list(clust_k10$cluster),FUN=mean)
# append cluster assignment
X_k10 = data.frame(X_scaled, clust_k10$cluster)
```

```{r}
cluster_means[ apply(cluster_means, 1, max) > 5, ]
```

```{r}
cluster_means[cluster_means$personal_fitness > 1.25, ]
```

Visualizations:
```{r}
kmeans_pl = fviz_cluster(clust_k10, data = X_scaled)

ggplotly(kmeans_pl)
```



Hierachical clustering using pre-processed data:
```{r}
# Center/scale the data
X_scaled = scale(X, center=TRUE, scale=TRUE) 

# Form a pairwise distance matrix using the dist function
market_distance_matrix = dist(X_scaled, method='euclidean')

# Run hierarchical clustering
hier_market = hclust(market_distance_matrix, method='complete')

# Plot the dendrogram
#plot(hier_market, cex=0.8)

# Cut the tree into 10 clusters
hcluster1 = cutree(hier_market, k=10)

```

```{r}
for (i in 1:10){
  len = length(which(data.frame(hcluster1)$hcluster1 == i))
  cat("Cluster", i, "has" , len, "members\n")
}
```

```{r}
## get cluster means 
#hclust_means = aggregate(X_scaled,by=list(hcluster1$cluster),FUN=mean)

for (i in 1:10){
  assign(paste('cluster', i, sep = ' '), colMeans(X_scaled[which(hcluster1 == i),]))
}
```

```{r}
#hclust_means = data.frame(X_scaled,hclust_assign)

##need to make df of means
```


```{r}
hclust_pl = fviz_cluster(list(data = X, cluster = hcluster1))

ggplotly(hclust_pl)
```



Run PCA and repeat kmeans:
```{r}
pc_market = prcomp(X, scale.=TRUE)

#summary(pc_market)
#plot(pc_market)

# Where do the individual points end up in PC space?
#biplot(pc_market)

loadings = pc_market$rotation
scores = pc_market$x

```

```{r}
pc_clust_k10 = kmeanspp(scores, k=10, nstart=25)

for (i in 1:10){
  len = length(which(data.frame(X_scaled, pc_clust_k10$cluster) == i))
  cat("Cluster", i, "has" , len, "members\n")
}
```

```{r}
## get cluster means 
pc_cluster_means = aggregate(X_scaled,by=list(pc_clust_k10$cluster),FUN=mean)

```

```{r}
pc_cluster_means[pc_cluster_means$personal_fitness > 1.25, ]
```

```{r}
pc_pl = fviz_cluster(pc_clust_k10, data = X_scaled)

ggplotly(pc_pl)
```


